Supplement to `\ourlongtitle'

### 2021-10-13 Final_version/glo-2020-proof-asymp.tex
Added final version

\begin{proof}[Proof of Proposition~\ref{prop:hoeffding:tau:marginal}]
We establish (i) by induction over $n$. Write, using Lemma~\ref{lem:expectation:incremental:marginal},
\begin{multline} \label{eq:hoeffding:decomp}
\frac{1}{\N} \sum_{i = 1}^\N \ewght{n + 1}{i} \{\tstat[i]{n + 1} f_{n + 1}(\epart{n + 1}{i}) + \ftd{n + 1}(\epart{n + 1}{i})\} - \frac{\postmod{n} \ukmod{n} \{ \tstatmod{n + 1} h_{n + 1} f_{n + 1} + \ftd{n + 1} \}}{\postmod{n} \adjfuncforward{n}}  \\
= \frac{1}{\N} \sum_{i = 1}^\N \ewght{n + 1}{i} \{\tstat[i]{n + 1} f_{n + 1}(\epart{n + 1}{i}) + \ftd{n + 1}(\epart{n + 1}{i})\} - \pE \left[ \ewght{n + 1}{1} \{\tstat[1]{n + 1} f_{n + 1}(\epart{n + 1}{1}) + \ftd{n + 1}(\epart{n + 1}{1})\} \cond \calF{n}{\N} \right] \\
+ \sum_{j = 1}^\N 
\frac{\ewght{n}{j} \adjfuncforward{n}(\epart{n}{j})}{\sum_{j' = 1}^\N \ewght{n}{j'} \adjfuncforward{n}(\epart{n}{j'})} \{ \tstat[j]{n} \prop{n} (\wgtfuncmod{n} f_{n + 1})(\epart{n}{j}) + \prop{n}(\wgtfuncmod{n} \addf{n} f_{n + 1} + \wgtfuncmod{n} \ftd{n + 1})(\epart{n}{j})\} \\
- \frac{\postmod{n - 1} \ukmod{n - 1}\{\tstatmod{n} \ukmod{n} f_{n + 1} + \ukmod{n} (\addf{n} f_{n + 1} + \ftd{n + 1})\}}{\postmod{n - 1} \ukmod{n - 1} \adjfuncforward{n}}. 
\end{multline}
Since for all $i$, 
$$
|\ewght{n + 1}{i} \{\tstat[i]{n + 1} f_{n + 1}(\epart{n + 1}{i}) + \ftd{n + 1}(\epart{n + 1}{i})\}| \leq \| \wgtfunc{n} \|_\infty (\| h_{n + 1} \|_\infty \| f_{n + 1} \|_\infty + \| \ftd{n + 1} \|_\infty), 
$$
the Hoeffding inequality for conditional expectations provides constants $(d, \tilde{d}) \in \rsetpos^2$ such that for all $\delta > 0$, 
\begin{multline*}
\pP \left( \left| \frac{1}{\N} \sum_{i = 1}^\N \ewght{n + 1}{i} \{\tstat[i]{n + 1} f_{n + 1}(\epart{n + 1}{i}) + \ftd{n + 1}(\epart{n + 1}{i})\} \right. \right. \\
\left. \left.  - \pE \left[ \ewght{n + 1}{1} \{\tstat[1]{n + 1} f_{n + 1}(\epart{n + 1}{1}) + \ftd{n + 1}(\epart{n + 1}{1})\} \cond \calF{n}{\N} \right] \vphantom{\sum_{i = 1}^\N} \right| \geq \delta \right) \leq d \exp \left( - \tilde{d} N \delta^2 \right).  
\end{multline*}
In addition, since $\adjfuncforward{n}$, $\adjfuncforward{n} \prop{n} (\wgtfuncmod{n} f_{n + 1})$, and $\adjfuncforward{n} \prop{n}(\wgtfuncmod{n} \addf{n} f_{n + 1} + \wgtfuncmod{n} \ftd{n + 1})$ all belong to $\bmf{\Xfd_n}$, there exist, by the induction hypothesis and \cite[Lemma~4]{douc:garivier:moulines:olsson:2010}, $(d', \tilde{d}') \in \rsetpos^2$ such that for all $\delta > 0$, 
\begin{multline*}
\pP \left( \left|  \sum_{j = 1}^\N 
\frac{\ewght{n}{j} \adjfuncforward{n}(\epart{n}{j})}{\sum_{j' = 1}^\N \ewght{n}{j'} \adjfuncforward{n}(\epart{n}{j'})} \{ \tstat[j]{n} \prop{n} (\wgtfuncmod{n} f_{n + 1})(\epart{n}{j}) + \prop{n}(\wgtfuncmod{n} \addf{n} f_{n + 1} + \wgtfuncmod{n} \ftd{n + 1})(\epart{n}{j})\} \right. \right. \\
\left. \left. - \frac{\postmod{n - 1} \ukmod{n - 1}\{\tstatmod{n} \ukmod{n} f_{n + 1} + \ukmod{n} (\addf{n} f_{n + 1} + \ftd{n + 1})\}}{\postmod{n - 1} \ukmod{n - 1} \adjfuncforward{n}} \right| \geq \delta \right) \leq d' \exp \left( - \tilde{d}' N \delta^2 \right).  
\end{multline*}
Combining the previous two inequalities completes the proof of (i) at time step $n + 1$. The base case $n = 1$ is established using again the decomposition \eqref{eq:hoeffding:decomp}, the standard Hoeffding inequality for independent and identically distributed variables, \cite[Lemma~4]{douc:garivier:moulines:olsson:2010}, and the fact that $h_0 \equiv 0$.

@@

We now turn our focus to the proof of Theorem~\ref{cor:clt:pseudo:marginal:paris}, which, just like Theorem~\ref{cor:hoeffding:tau:marginal}, will be established via the following more general result.

\begin{theorem}
\label{prop:clt:pseudo:marginal:paris}
Assume~\hypref[assum:biased:estimate]{assum:bound:filter:pseudomarginal}. Then for all $n \in \nset$, $\K \in \nsetpos$, and $(f_n, \ftd{n}) \in \bmf{\Xfd_n}^2$,
$$
 \sqrt{\N} \left( \sum_{i = 1}^\N \frac{\ewght{n}{i}}{\sumwght{n}} \{\tstat[i]{n} f_n(\epart{n}{i}) + \ftd{n}(\epart{n}{i}) \} - \postmod{n}(\tstatmod{n} h_n f_n + \ftd{n}) \right) 
  \dlim \sigma_n(h_n, f_n, \ftd{n}) Z,
$$
where $Z$ is standard normally distributed and 
\begin{equation} \label{eq:non-recursive:as:var:affine}
\sigma^2_n (h_n, f_n, \ftd{n}) \eqdef \frac{\chi \{ \initwgtfunc
\retrokmodmodnorm_{0, n} (h_n f_n + \ftd{n}) \}^2}{(\chi \ukmod{0, n - 1} \1_{\Xset^n})^2} 
+ \sigma^2_n \langle (\wgtfuncmod{\ell})_{\ell = 0}^{n - 1} \rangle (h_n, f_n, \ftd{n}) 
+ \sigma^2_n \langle (\ukestvar{\ell})_{\ell = 0}^{n - 1} \rangle (h_n, f_n, \ftd{n}) 
\end{equation}
and
\begin{multline} \label{eq:partial:variance}
\sigma^2_n \langle (\varphi_\ell)_{\ell = 0}^{n - 1} \rangle (h_n, f_n, \ftd{n}) 
\eqdef \sum_{m = 0}^{n - 1} \frac{\postmod{m} \adjfuncforward{m} \postmod{m} \ukmod{m} \{ \varphi_m [\retrokmodmodnorm_{m + 1, n} (h_n f_n + \ftd{n})]^2
\}}{(\postmod{m} \ukmod{m, n - 1} \1_{\Xset_n})^2} \\
+ \sum_{m = 0}^{n - 1} \sum_{\ell = 0}^m \frac{\postmod{m} \adjfuncforward{m} \postmod{\ell} \ukmod{\ell} \{\bkmod{\ell}(\tstatmod{\ell} h_{\ell} + \addf{\ell} - \tstatmod{\ell + 1} h_{\ell +1})^2 \ukmod{\ell + 1, m}( \bkmod{m} \varphi_m [\ukmod{m + 1, n - 1} f_n]^2
)\}}{\K^{m - \ell + 1} (\postmod{\ell} \ukmod{\ell, m - 1} \1_{\Xset_m})(\postmod{m} \ukmod{m, n - 1} \1_{\Xset_n})^2}
\end{multline}
for any sequence $(\varphi_\ell)_{\ell \in \nset}$ of measurable functions $\varphi_n : \Xset_n \times \Xset_{n + 1} \to \rsetnn$. 
\end{theorem}

As previously, applying Theorem~\ref{prop:clt:pseudo:marginal:paris} with $f_n \equiv 1$ and $\ftd{n} \equiv 0$ yields immediately Theorem~\ref{cor:hoeffding:tau:marginal}. Before proving Theorem~\ref{prop:clt:pseudo:marginal:paris} we establish some preparatory lemmas, where the first is easily obtained by combining Lemma~\ref{lem:reversibility}(i) (applied to the skew modeled formed by $\chi$ and $(\ukmod{n})_{n \in \nset}$) and \eqref{eq:forward:smoothing}.

\begin{lemma} \label{eq:critical:identity}
Assume \hypref{assum:biased:estimate}. Then for all $n \in \nset$ and $(f_{n + 1}, \ftd{n + 1}) \in \bmf{\Xfd_{n + 1}}^2$, 
$$
\postmod{n} \ukmod{n}(\tstatmod{n + 1} h_{n + 1} f_{n + 1} + \ftd{n + 1}) = \postmod{n}\{\tstatmod{n} \ukmod{n} f_{n + 1} + \ukmod{n} (\addf{n} f_{n + 1} + \ftd{n + 1})\}. 
$$
\end{lemma}

\begin{lemma}
\label{lem:hoeffding:tau:square}
Assume \hypref[assum:biased:estimate]{assum:bound:filter:pseudomarginal}. Then for all $n \in \nset$, $f \in \bmf{\Xfd_n}$, and $\K \in \nsetpos$, 
\[
\sum_{i = 1}^\N \frac{\ewght{n}{i}}{\sumwght{n}}(\tstat[i]{n})^2 f(\epart{n}{i}) \pplim \postmod{n}([\tstatmod{n} h_n]^2 f) + \eta_n f, 
\]
where the measures $(\eta_m)_{m \in \nsetpos}$ are defined recursively as 
\[
\eta_{m + 1} f = \K^{-1} \frac{\eta_m \ukmod{m} f +  \postmod{m} \ukmod{m} \{ \bkmod{m}(\tstatmod{m} h_m + \addf{m} - \tstatmod{m + 1} h_{m + 1})^2 f \}}{\postmod{m} \ukmod{m} \1_{\Xset_{m + 1}}}, \quad m \in \nset, 
\]
with $\eta_0 \equiv 0$. 
\end{lemma}

Following \cite[Lemma~13]{olsson:westerborn:2017}, the measures $(\eta_m)_{m \in \nsetpos}$ may be expressed non-recursively as
\begin{equation}
\label{eq:def:etak}
\eta_n f = \sum_{m = 0}^{n - 1} \K^{m - n} \frac{\postmod{m} \ukmod{m} \{ \bkmod{m}(\tstatmod{m} h_m + \addf{m} - \tstatmod{m + 1} h_{m + 1})^2 \ukmod{m + 1, n - 1} f \}}{\postmod{m} \ukmod{m, n - 1} \1_{\Xset_n}}.
\end{equation}

\begin{proof}[Proof of Lemma~\ref{lem:hoeffding:tau:square}]
We proceed by induction over $n$. The base case $n = 0$ is a trivial consequence of the fact that $\tstat{0} h_0 = 0$ and $\tstat[i]{0} = 0$ for all $i \in \intvect{1}{\N}$. Thus, we assume that the result holds true for some $n \in \nset$ and write
\[
\sum_{i=1}^\N \frac{\ewght{n+1}{i}}{\sumwght{n+1}}(\tstat[i]{n+1})^2 f(\epart{n+1}{i}) = \frac{a_\N}{b_\N},
\]
where 
$$
a_\N \eqdef \frac{1}{\N}\sum_{i=1}^N\ewght{n+1}{i}(\tstat[i]{n+1})^2 f(\epart{n+1}{i}), \quad b_\N \eqdef \frac{1}{\N} \sum_{i=1}^N \ewght{n+1}{i}
$$

We first establish the convergence in probability of $(a_\N)_{\N \in \nsetpos}$. Using again Hoeffding's inequality for conditional expectations and the fact that the variables $\ewght{n+1}{i}(\tstat[i]{n + 1})^2 f_{n + 1}(\epart{n + 1}{i})$, $i \in \intvect{1}{\N}$, are all bounded by $\| \wgtfunc{n} \|_\infty \| h_{n + 1} \|^2_\infty \| f_{n + 1}\|_\infty$ and conditionally i.i.d. given $\calF{n}{\N}$ for all $i$, we obtain for all $\epsilon > 0$,  
\[
\pP\left(\left| a_\N - \pE[a_\N \mid \calF{n}{\N}] \right| \geq \epsilon \right) \leq \exp \left( - \frac{2 \N \epsilon^2}{\| \wgtfunc{n} \|_\infty \| h_{n + 1} \|^2_\infty \| f_{n+1} \|_\infty} \right). 
\]
It is hence enough to consider the limit in probability of $\pE[a_\N \mid \calF{n}{\N}]$ as $\N$ tends to infinity. For this purpose, write, using \eqref{eq:expected:weight:given:G}, 
\begin{align*}
\pE[a_\N \mid \calF{n}{\N}] &= \pE \left[ \ewght{n + 1}{1}(\tstat[1]{n + 1})^2 f_{n+1}(\epart{n + 1}{1}) \mid  \calF{n}{\N} \right] \\ 
&= \pE\left[\wgtfuncmod{n}(\epart{n}{\ind{n + 1}{1}}, \epart{n + 1}{1}) \pE \left[(\tstat[1]{n + 1})^2 \cond \calG{n + 1}{\N}\right] f_{n + 1}(\epart{n + 1}{1}) \cond \calF{n}{\N} \right] \\
&= a_\N' + a_\N'',
\end{align*}
where
\begin{align*}
a_\N' &\eqdef \K^{-1}\pE\left[ \wgtfuncmod{n}(\epart{n}{\ind{n + 1}{1}}, \epart{n + 1}{1}) f_{n + 1}(\epart{n + 1}{1}) \pE \left[ \left( \tstat[J_{n + 1}^{(1,1)}]{n} + \addf{n}(\epart{n}{J_{n + 1}^{(1,1)}}, \epart{n + 1}{1}) \right)^2 \cond \calG{n + 1}{\N}\right] \cond \calF{n}{\N} \right], \\
a_\N'' &\eqdef (\K - 1)\K^{-1} \pE \left[ \wgtfuncmod{n}(\epart{n}{\ind{n + 1}{1}}, \epart{n + 1}{1}) f_{n + 1}(\epart{n + 1}{1}) \pE \left[\tstat[J_{n + 1}^{(1,1)}]{n} + \addf{n}(\epart{n}{J_{n + 1}^{(1,1)}}, \epart{n + 1}{1}) \cond \calG{n + 1}{\N}\right]^2 \cond \calF{n}{\N} \right]. 
\end{align*}
Here the first term is given by
\begin{align}
a_\N' &= \K^{-1} \sum_{i = 1}^\N \frac{\ewght{n}{i}\adjfuncforward{n}(\epart{n}{i})}{\sum_{i' = 1}^{\N} \ewght{n}{i'} \adjfuncforward{n}(\epart{n}{i'})} \int \wgtfuncmod{n}(\epart{n}{i}, x) f_{n + 1}(x) \nonumber \\
&\hspace{3cm}\times \sum_{j = 1}^\N \frac{\ewght{n}{j} \udmod{n}(\epart{n}{j}, x)}{\sum_{j' = 1}^{\N} \ewght{n}{j'} \udmod{n}(\epart{n}{j'}, x)} \left(\tstat[j]{n} + \addf{n}(\epart{n}{j}, x) \right)^2  \kissforward{n}{n}(\epart{n}{i},x) \, \mu(\rmd x), \nonumber \\
&= (\K \post[\N]{n}\adjfuncforward{n})^{-1} \sum_{j = 1}^\N \frac{\ewght{n}{j}}{\sumwght{n}} \int f_{n + 1}(x) \left( \tstat[j]{n} + \addf{n}(\epart{n}{j}, x) \right)^2 \, \ukmod{n}(\epart{n}{j}, \rmd x). \label{eq:a':1}
\end{align}
Thus, by the induction hypothesis and Proposition~\ref{prop:hoeffding:tau:marginal}, $a_\N'$ tends in probability to 
\begin{multline*}
(\K \postmod{n}\adjfuncforward{n})^{-1} \left(\eta_n \ukmod{n} f_{n + 1} + \postmod{n}\{(\tstatmod{n} h_n)^2 \ukmod{n} f_{n+1}\} + \postmod{n} \ukmod{n}(f_{n + 1} \addf{n}^2)  + 2 \postmod{n}\{\tstatmod{n} h_n \ukmod{n}(f_{n + 1} \addf{n})\} \right) \\
= (\K \postmod{n} \adjfuncforward{n})^{-1} \left(\eta_n \ukmod{n} f_{n + 1} + \postmod{n} \ukmod{n} \{(\tstatmod{n} h_n +\addf{n})^2 f_{n + 1} \} \right). 
\end{multline*}

@@

which verifies condition (i) in \cite[Theorem~A.3]{douc:moulines:2008}. Now, with the conditional relative weight variance $\ukestvar{n}$ defined as in \eqref{eq:def:ukestvar} it holds that  

@@

+ \frac{\postmod{n} \adjfuncforward{n}}{(\postmod{n} \ukmod{n} \1_{\Xset_{n + 1}})^2} \sum_{m = 0}^n \frac{\postmod{m} \ukmod{m}\{\bkmod{m}(\tstatmod{m} h_m + \addf{m} - \tstatmod{m + 1} h_{m + 1})^2 \ukmod{m + 1, n}(f_{n + 1}^2 \bkmod{n} \wgtfuncmod{n})\}}{\K^{n - m + 1} \postmod{m} \ukmod{m, n - 1} \1_{\Xset_n}} \\

>>>>

\begin{proof}[Proof of Proposition~\ref{prop:hoeffding:tau:marginal}]
We establish (i) by induction over $n$. Write, using Lemma~\ref{lem:expectation:incremental:marginal},
\begin{multline} \label{eq:hoeffding:decomp}
\frac{1}{\N} \sum_{i = 1}^\N \ewght{n + 1}{i} \{\tstat[i]{n + 1} f_{n + 1}(\epart{n + 1}{i}) + \ftd{n + 1}(\epart{n + 1}{i})\} - \frac{\postmod{n} \ukmod{n} (\tstatmod{n + 1} h_{n + 1} f_{n + 1} + \ftd{n + 1})}{\postmod{n} \adjfuncforward{n}}  \\
= \frac{1}{\N} \sum_{i = 1}^\N \ewght{n + 1}{i} \{\tstat[i]{n + 1} f_{n + 1}(\epart{n + 1}{i}) + \ftd{n + 1}(\epart{n + 1}{i})\} - \pE \left[ \ewght{n + 1}{1} \{\tstat[1]{n + 1} f_{n + 1}(\epart{n + 1}{1}) + \ftd{n + 1}(\epart{n + 1}{1})\} \cond \calF{n}{\N} \right] \\
+ \sum_{j = 1}^\N 
\frac{\ewght{n}{j} \adjfuncforward{n}(\epart{n}{j})}{\sum_{j' = 1}^\N \ewght{n}{j'} \adjfuncforward{n}(\epart{n}{j'})} \{ \tstat[j]{n} \prop{n} (\wgtfuncmod{n} f_{n + 1})(\epart{n}{j}) + \prop{n}(\wgtfuncmod{n} \addf{n} f_{n + 1} + \wgtfuncmod{n} \ftd{n + 1})(\epart{n}{j})\} \\
- \frac{\postmod{n - 1} \ukmod{n - 1}\{\tstatmod{n} h_n \ukmod{n} f_{n + 1} + \ukmod{n} (\addf{n} f_{n + 1} + \ftd{n + 1})\}}{\postmod{n - 1} \ukmod{n - 1} \adjfuncforward{n}}. 
\end{multline}
Since for all $i$, 
$$
|\ewght{n + 1}{i} \{\tstat[i]{n + 1} f_{n + 1}(\epart{n + 1}{i}) + \ftd{n + 1}(\epart{n + 1}{i})\}| \leq \| \wgtfunc{n} \|_\infty (\| h_{n + 1} \|_\infty \| f_{n + 1} \|_\infty + \| \ftd{n + 1} \|_\infty), 
$$
the Hoeffding inequality for conditional expectations provides constants $(d, \tilde{d}) \in \rsetpos^2$ such that for all $\delta > 0$, 
\begin{multline*}
\pP \left( \left| \frac{1}{\N} \sum_{i = 1}^\N \ewght{n + 1}{i} \{\tstat[i]{n + 1} f_{n + 1}(\epart{n + 1}{i}) + \ftd{n + 1}(\epart{n + 1}{i})\} \right. \right. \\
\left. \left.  - \pE \left[ \ewght{n + 1}{1} \{\tstat[1]{n + 1} f_{n + 1}(\epart{n + 1}{1}) + \ftd{n + 1}(\epart{n + 1}{1})\} \cond \calF{n}{\N} \right] \vphantom{\sum_{i = 1}^\N} \right| \geq \delta \right) \leq d \exp \left( - \tilde{d} N \delta^2 \right).  
\end{multline*}
In addition, since $\adjfuncforward{n}$, $\adjfuncforward{n} \prop{n} (\wgtfuncmod{n} f_{n + 1})$, and $\adjfuncforward{n} \prop{n}(\wgtfuncmod{n} \addf{n} f_{n + 1} + \wgtfuncmod{n} \ftd{n + 1})$ all belong to $\bmf{\Xfd_n}$, there exist, by the induction hypothesis and \cite[Lemma~4]{douc:garivier:moulines:olsson:2010}, $(d', \tilde{d}') \in \rsetpos^2$ such that for all $\delta > 0$, 
\begin{multline*}
\pP \left( \left|  \sum_{j = 1}^\N 
\frac{\ewght{n}{j} \adjfuncforward{n}(\epart{n}{j})}{\sum_{j' = 1}^\N \ewght{n}{j'} \adjfuncforward{n}(\epart{n}{j'})} \{ \tstat[j]{n} \prop{n} (\wgtfuncmod{n} f_{n + 1})(\epart{n}{j}) + \prop{n}(\wgtfuncmod{n} \addf{n} f_{n + 1} + \wgtfuncmod{n} \ftd{n + 1})(\epart{n}{j})\} \right. \right. \\
\left. \left. - \frac{\postmod{n - 1} \ukmod{n - 1}\{\tstatmod{n} h_n \ukmod{n} f_{n + 1} + \ukmod{n} (\addf{n} f_{n + 1} + \ftd{n + 1})\}}{\postmod{n - 1} \ukmod{n - 1} \adjfuncforward{n}} \right| \geq \delta \right) \leq d' \exp \left( - \tilde{d}' N \delta^2 \right).  
\end{multline*}
Combining the previous two inequalities completes the proof of (i) at time step $n + 1$. The base case $n = 1$ is established using again the decomposition \eqref{eq:hoeffding:decomp}, the standard Hoeffding inequality for independent and identically distributed variables, \cite[Lemma~4]{douc:garivier:moulines:olsson:2010}, and the fact that $h_0 \equiv 0$.

@@

We now turn our focus to the proof of Theorem~\ref{cor:clt:pseudo:marginal:paris}, which, like Theorem~\ref{cor:hoeffding:tau:marginal}, will be established via the following more general result.

\begin{theorem}
\label{prop:clt:pseudo:marginal:paris}
Assume~\hypref[assum:biased:estimate]{assum:bound:filter:pseudomarginal}. Then for all $n \in \nset$, $\K \in \nsetpos$, and $(f_n, \ftd{n}) \in \bmf{\Xfd_n}^2$, as $\N \to \infty$, 
$$
 \sqrt{\N} \left( \sum_{i = 1}^\N \frac{\ewght{n}{i}}{\sumwght{n}} \{\tstat[i]{n} f_n(\epart{n}{i}) + \ftd{n}(\epart{n}{i}) \} - \postmod{n}(\tstatmod{n} h_n f_n + \ftd{n}) \right) 
  \dlim \sigma_n(h_n, f_n, \ftd{n}) Z,
$$
where $Z$ is standard normally distributed and 
\begin{equation} \label{eq:non-recursive:as:var:affine}
\sigma^2_n (h_n, f_n, \ftd{n}) \eqdef \frac{\chi \{ \initwgtfunc
\retrokmodmodnorm_{0, n} (h_n f_n + \ftd{n}) \}^2}{(\chi \ukmod{0, n - 1} \1_{\Xset^n})^2} 
+ \sigma^2_n \langle (\wgtfuncmod{\ell})_{\ell = 0}^{n - 1} \rangle (h_n, f_n, \ftd{n}) 
+ \sigma^2_n \langle (\ukestvar{\ell})_{\ell = 0}^{n - 1} \rangle (h_n, f_n, \ftd{n}) 
\end{equation}
with
\begin{multline} \label{eq:partial:variance}
\sigma^2_n \langle (\varphi_\ell)_{\ell = 0}^{n - 1} \rangle (h_n, f_n, \ftd{n}) 
\eqdef \sum_{m = 0}^{n - 1} \frac{\postmod{m} \adjfuncforward{m} \postmod{m} \ukmod{m} \{ \varphi_m [\retrokmodmodnorm_{m + 1, n} (h_n f_n + \ftd{n})]^2
\}}{(\postmod{m} \ukmod{m, n - 1} \1_{\Xset^n})^2} \\
+ \sum_{m = 0}^{n - 1} \sum_{\ell = 0}^m \frac{\postmod{m} \adjfuncforward{m} \postmod{\ell} \ukmod{\ell} \{\bkmod{\ell}(\tstatmod{\ell} h_{\ell} + \addf{\ell} - \tstatmod{\ell + 1} h_{\ell +1})^2 \ukmod{\ell + 1, m}( \bkmod{m} \varphi_m [\ukmod{m + 1, n - 1} f_n]^2
)\}}{\K^{m - \ell + 1} (\postmod{\ell} \ukmod{\ell, m - 1} \1_{\Xset^m})(\postmod{m} \ukmod{m, n - 1} \1_{\Xset^n})^2}
\end{multline}
for any sequence $(\varphi_\ell)_{\ell \in \nset}$ of measurable functions $\varphi_n : \Xset_n \times \Xset_{n + 1} \to \rsetnn$. 
\end{theorem}

As previously, applying Theorem~\ref{prop:clt:pseudo:marginal:paris} with $f_n \equiv 1$ and $\ftd{n} \equiv 0$ yields immediately Theorem~\ref{cor:hoeffding:tau:marginal}. Before proving Theorem~\ref{prop:clt:pseudo:marginal:paris} we establish some preparatory lemmas, where the first is easily obtained by combining Lemma~\ref{lem:reversibility}(i) (applied to the skew model formed by $\chi$ and $(\ukmod{n})_{n \in \nset}$) and \eqref{eq:forward:smoothing}.

\begin{lemma} \label{eq:critical:identity}
Assume \hypref{assum:biased:estimate}. Then for all $n \in \nset$ and $(f_{n + 1}, \ftd{n + 1}) \in \bmf{\Xfd_{n + 1}}^2$, 
$$
\postmod{n} \ukmod{n}(\tstatmod{n + 1} h_{n + 1} f_{n + 1} + \ftd{n + 1}) = \postmod{n}\{\tstatmod{n} h_n \ukmod{n} f_{n + 1} + \ukmod{n} (\addf{n} f_{n + 1} + \ftd{n + 1})\}. 
$$
\end{lemma}

\begin{lemma}
\label{lem:hoeffding:tau:square}
Assume \hypref[assum:biased:estimate]{assum:bound:filter:pseudomarginal}. Then for all $n \in \nset$, $f \in \bmf{\Xfd_n}$, and $\K \in \nsetpos$, as $\N \to \infty$, 
\[
\sum_{i = 1}^\N \frac{\ewght{n}{i}}{\sumwght{n}}(\tstat[i]{n})^2 f(\epart{n}{i}) \pplim \postmod{n}([\tstatmod{n} h_n]^2 f) + \eta_n f, 
\]
where the measures $(\eta_m)_{m \in \nsetpos}$ are defined recursively as 
\[
\eta_{m + 1} f = \K^{-1} \frac{\eta_m \ukmod{m} f +  \postmod{m} \ukmod{m} \{ \bkmod{m}(\tstatmod{m} h_m + \addf{m} - \tstatmod{m + 1} h_{m + 1})^2 f \}}{\postmod{m} \ukmod{m} \1_{\Xset_{m + 1}}}, \quad m \in \nset, 
\]
with $\eta_0 \equiv 0$. 
\end{lemma}

Following \cite[Lemma~13]{olsson:westerborn:2017}, the measures $(\eta_m)_{m \in \nsetpos}$ may be expressed non-recursively as
\begin{equation}
\label{eq:def:etak}
\eta_n f = \sum_{m = 0}^{n - 1} \K^{m - n} \frac{\postmod{m} \ukmod{m} \{ \bkmod{m}(\tstatmod{m} h_m + \addf{m} - \tstatmod{m + 1} h_{m + 1})^2 \ukmod{m + 1, n - 1} f \}}{\postmod{m} \ukmod{m, n - 1} \1_{\Xset^n}}.
\end{equation}

\begin{proof}[Proof of Lemma~\ref{lem:hoeffding:tau:square}]
We proceed by induction over $n$. The base case $n = 0$ is a trivial consequence of the fact that $\tstat{0} h_0 = 0$ and $\tstat[i]{0} = 0$ for all $i \in \intvect{1}{\N}$. Thus, we assume that the result holds true for some $n \in \nset$ and write
\[
\sum_{i=1}^\N \frac{\ewght{n+1}{i}}{\sumwght{n+1}}(\tstat[i]{n+1})^2 f(\epart{n+1}{i}) = \frac{a_\N}{b_\N},
\]
where 
$$
a_\N \eqdef \frac{1}{\N}\sum_{i=1}^N\ewght{n+1}{i}(\tstat[i]{n+1})^2 f(\epart{n+1}{i}), \quad b_\N \eqdef \frac{1}{\N} \sum_{i=1}^N \ewght{n+1}{i}. 
$$

We first establish the convergence in probability of $(a_\N)_{\N \in \nsetpos}$. Using again Hoeffding's inequality for conditional expectations and the fact that the variables $\ewght{n+1}{i}(\tstat[i]{n + 1})^2 f_{n + 1}(\epart{n + 1}{i})$, $i \in \intvect{1}{\N}$, are all bounded by $\| \wgtfunc{n} \|_\infty \| h_{n + 1} \|^2_\infty \| f_{n + 1}\|_\infty$ and conditionally i.i.d. given $\calF{n}{\N}$, we obtain for all $\epsilon > 0$,  
\[
\pP\left(\left| a_\N - \pE[a_\N \mid \calF{n}{\N}] \right| \geq \epsilon \right) \leq \exp \left( - \frac{2 \N \epsilon^2}{\| \wgtfunc{n} \|_\infty \| h_{n + 1} \|^2_\infty \| f_{n+1} \|_\infty} \right). 
\]
It is hence enough to consider the limit in probability of $\pE[a_\N \mid \calF{n}{\N}]$ as $\N$ tends to infinity. For this purpose, write, using \eqref{eq:expected:weight:given:G}, 
\begin{align*}
\pE[a_\N \mid \calF{n}{\N}] &= \pE \left[ \ewght{n + 1}{1}(\tstat[1]{n + 1})^2 f_{n+1}(\epart{n + 1}{1}) \mid  \calF{n}{\N} \right] \\ 
&= \pE\left[\wgtfuncmod{n}(\epart{n}{\ind{n + 1}{1}}, \epart{n + 1}{1}) \pE \left[(\tstat[1]{n + 1})^2 \cond \calG{n + 1}{\N}\right] f_{n + 1}(\epart{n + 1}{1}) \cond \calF{n}{\N} \right] \\
&= a_\N' + a_\N'',
\end{align*}
where
\begin{align*}
a_\N' &\eqdef \K^{-1}\pE\left[ \wgtfuncmod{n}(\epart{n}{\ind{n + 1}{1}}, \epart{n + 1}{1}) f_{n + 1}(\epart{n + 1}{1}) \pE \left[ \left( \tstat[J_{n + 1}^{(1,1)}]{n} + \addf{n}(\epart{n}{J_{n + 1}^{(1,1)}}, \epart{n + 1}{1}) \right)^2 \cond \calG{n + 1}{\N}\right] \cond \calF{n}{\N} \right], \\
a_\N'' &\eqdef (\K - 1)\K^{-1} \pE \left[ \wgtfuncmod{n}(\epart{n}{\ind{n + 1}{1}}, \epart{n + 1}{1}) f_{n + 1}(\epart{n + 1}{1}) \pE \left[\tstat[J_{n + 1}^{(1,1)}]{n} + \addf{n}(\epart{n}{J_{n + 1}^{(1,1)}}, \epart{n + 1}{1}) \cond \calG{n + 1}{\N}\right]^2 \cond \calF{n}{\N} \right]. 
\end{align*}
Here the first term is given by
\begin{align}
a_\N' &= \K^{-1} \sum_{i = 1}^\N \frac{\ewght{n}{i}\adjfuncforward{n}(\epart{n}{i})}{\sum_{i' = 1}^{\N} \ewght{n}{i'} \adjfuncforward{n}(\epart{n}{i'})} \int \wgtfuncmod{n}(\epart{n}{i}, x) f_{n + 1}(x) \nonumber \\
&\hspace{3cm}\times \sum_{j = 1}^\N \frac{\ewght{n}{j} \udmod{n}(\epart{n}{j}, x)}{\sum_{j' = 1}^{\N} \ewght{n}{j'} \udmod{n}(\epart{n}{j'}, x)} \left(\tstat[j]{n} + \addf{n}(\epart{n}{j}, x) \right)^2  \kissforward{n}{n}(\epart{n}{i},x) \, \mu(\rmd x), \nonumber \\
&= (\K \post[\N]{n}\adjfuncforward{n})^{-1} \sum_{j = 1}^\N \frac{\ewght{n}{j}}{\sumwght{n}} \int f_{n + 1}(x) \left( \tstat[j]{n} + \addf{n}(\epart{n}{j}, x) \right)^2 \, \ukmod{n}(\epart{n}{j}, \rmd x). \label{eq:a':1}
\end{align}
Thus, by the induction hypothesis and Proposition~\ref{prop:hoeffding:tau:marginal}, $a_\N'$ tends in probability to 
\begin{multline*}
(\K \postmod{n}\adjfuncforward{n})^{-1} \left(\eta_n \ukmod{n} f_{n + 1} + \postmod{n}\{(\tstatmod{n} h_n)^2 \ukmod{n} f_{n+1}\} + \postmod{n} \ukmod{n}(f_{n + 1} \addf{n}^2)  + 2 \postmod{n}\{\tstatmod{n} h_n \ukmod{n}(f_{n + 1} \addf{n})\} \right) \\
= (\K \postmod{n} \adjfuncforward{n})^{-1} \left(\eta_n \ukmod{n} f_{n + 1} + \postmod{n} \ukmod{n} \{(\tstatmod{n} h_n +\addf{n})^2 f_{n + 1} \} \right). 
\end{multline*}

@@

which verifies condition (i) in \cite[Theorem~A.3]{douc:moulines:2008}. Now, with the conditional relative weight variance $\ukestvar{n}$ defined in \eqref{eq:def:ukestvar} it holds that  

@@

+ \frac{\postmod{n} \adjfuncforward{n}}{(\postmod{n} \ukmod{n} \1_{\Xset_{n + 1}})^2} \sum_{m = 0}^n \frac{\postmod{m} \ukmod{m}\{\bkmod{m}(\tstatmod{m} h_m + \addf{m} - \tstatmod{m + 1} h_{m + 1})^2 \ukmod{m + 1, n}(f_{n + 1}^2 \bkmod{n} \wgtfuncmod{n})\}}{\K^{n - m + 1} \postmod{m} \ukmod{m, n - 1} \1_{\Xset^n}} \\
